\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in \proglang{R}

### 2014-12-15 papers/jss/response-to-reviewers.tex

\documentclass[10pt]{article}
\usepackage{url}
\usepackage{vmargin}
\setpapersize{USletter}
% left top right bottom -- headheight headsep footheight footskop
\setmarginsrb{1in}{1in}{1in}{0.5in}{0pt}{0mm}{10pt}{0.5in}
\usepackage{charter}

\setlength{\parskip}{1ex plus1ex minus1ex}
\setlength{\parindent}{0pt}

\newcommand{\proglang}[1]{\textsf{#1}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}

\newcommand{\pointRaised}[2]{\smallskip %\hrule 
  \textsl{{\fontseries{b}\selectfont #1}: #2}\newline}
\newcommand{\simplePointRaised}[1]{\bigskip \hrule\textsl{#1} }
\newcommand{\reply}[1]{\textbf{Reply}:\ #1 \smallskip } %\hrule \smallskip}

\begin{document}

\author{Dirk Eddelbuettel\\Debian Project \and 
        Murray Stokely\\Google, Inc \and
        Jeroen Ooms\\UCLA}
\title{Submission JSS 1313: \\ Response to Reviewers' Comments}
\maketitle 
\thispagestyle{empty}

Thank you for reviewing our manuscript, and for giving us an opportunity to
rewrite, extend and and tighten both the paper and the underlying package.

\smallskip
We truly appreciate the comments and suggestions. Below, we have regrouped the sets
of comments, and have provided detailed point-by-point replies.
%
We hope that this satisfies the request for changes necessary to proceed with
the publication of the revised and updated manuscript, along with the revised
and updated package (which was recently resubmitted to CRAN as version 0.4.2).

\section*{Response to Reviewer \#1}

\pointRaised{Comment 1}{Overall, I think this is a strong paper. Cross-language communication
  is a challenging problem, and good solutions for R are important to
  establish R as a well-behaved member of a data analysis pipeline. The
  paper is well written, and I recommend that it be accepted subject to
  the suggestions below.}
\reply{Thank you. We are providing a point-by-point reply below.}

\subsubsection*{More big picture, less details}

\pointRaised{Comment 2}{Overall, I think the paper provides too much detail on
  relatively unimportant topics and not enough on the reasoning behind
  important design decisions. I think you could comfortably reduce the paper
  by 5-10 pages, referring the interested reader to the documentation for
  more detail.}
\reply{The paper was rewritten throughout and is now much tighter at just 23 pages.}

\pointRaised{Comment 3}{I'd recommend shrinking section 3 to ~2 pages, and removing the
  subheadings. This section should quickly orient the reader to the
  RProtobuf API so they understand the big picture before learning more
  details in the subsequent sections. I'd recommend picking one OO style
  and sticking to it in this section - two is confusing.}
\reply{We followed this recommendation and reduced section 3 to about 2 1/2 pages.}

\pointRaised{Comment 3}{Section 4 dives into the details without giving a good overview and
  motivation. Why use S4 and not RC? How are the objects made mutable?
  Why do you provide both generic function and message passing OO
  styles? What does \$ do in this context? What the heck is a
  pseudo-method? Spend more time on those big issues rather than
  describing each class in detail. Reduce class descriptions to a
  bulleted list giving a high-level overview, then encourage the reader
  to refer to the documentation for further details. Similarly, Tables
  3-5 belong in the documentation, not in a vignette/paper.}
\reply{Done. TO BE EXPANDED}

\pointRaised{Comment 4}{Section 7 is weak. I think the important message is that RProtobuf is
  being used in practice at large scale for for large data, and is
  useful for communicating between R and Python. How can you make that
  message stronger while avoiding (for the purposes of this paper) the
  relatively unimportant details of the map-reduce setup?}
\reply{TBD}

\subsubsection*{R to/from Protobuf translation}

\pointRaised{Comment 5}{The discussion of R to/from Protobuf could be improved. Table 9 would be
  much simpler if instead of Message, you provided a "vectorised"
  Messages class (this would also make the interface more consistent and
  hence the package easier to use).}
\reply{TBD}

\pointRaised{Comment 6}{Along these lines, I think it would make sense to combine sections 5
  and 6 and discuss translation challenges in both direction
  simultaneously. At the minimum, add the equivalent for Table 9 that
  shows how important R classes are converted to their protobuf
  equivalents.}
\reply{TBD}

\pointRaised{Comment 7}{You should discuss how missing values are handled for strings and
  integers, and why enums are not equivalent to factors. I think you
  could make explicit how coercion of factors, dates, times and matrices
  occurs, and the implications of this on sharing data structures
  between programming languages. For example, how do you share date/time
  data between R and python using RProtoBuf?}
\reply{TBD}

\pointRaised{Comment 8}{Table 10 is dying to be a plot, and a natural companion would be to
  show how long it takes to serialise data frames using both RProtoBuf
  and R's native serialisation. Is there a performance penalty to using
  protobufs?}
\reply{TBD}

\subsubsection*{RObjectTables magic}

\pointRaised{Comment 9}{The use of RObjectTables magic makes me uneasy. It doesn't seem like a
  good fit for an infrastructure package and it's not clear what
  advantages it has over explicitly loading a protobuf definition into
  an object.}
\reply{TBD}

\pointRaised{Comment 10}{Using global state makes understanding code much harder. In Table 1,
  it's not obvious where \texttt{tutorial.Person} comes from. Is it loaded by
  default by RProtobuf? This need some explanation. In Section 7, what
  does \texttt{readProtoFiles()} do? Why does \texttt{RProtobuf} need to be attached
  as well as \texttt{HistogramTools}? This needs more explanation, and a
  comment on the implications of this approach on CRAN packages and
  namespaces.}
\reply{TBD}

\pointRaised{Comment 11}{
  I'd prefer you eliminate this magic from the magic, but failing that,
  you need a good explanation of why.}
\reply{TBD}

\subsubsection*{Code comments}

\pointRaised{Comment 12}{Using \texttt{file.create()} to determine the absolute path seems like a bad idea.}
\reply{TBD}

\subsubsection*{Minor niggles}

\pointRaised{Comment 13}{Don't refer to the message passing style of OO as traditional.}
\reply{TBD}

\pointRaised{Comment 14}{In Section 3.4, if messages isn't a vectorised class, the default
   print method should use \texttt{cat()} to eliminate the confusing \texttt{[1]}.}
\reply{TBD}

\pointRaised{Comment 15}{The REXP definition would have been better defined using an enum that
   matches R's SEXPTYPE "enum". But I guess that ship has sailed.}
\reply{TBD}

\pointRaised{Comment 16}{Why does \texttt{serialize\_pb(CO2, NULL)} fail silently? Shouldn't it at least
   warn that the serialization is partial?}
\reply{TBD}

\section*{Response to Reviewer \#2}

\pointRaised{Comment 1}{The paper gives an overview of the RProtoBuf package which implements an 
  R interface to the Protocol Buffers library for an efficient 
  serialization of objects. The paper is well written and easy to read. 
  Introductory code is clear and the package provides objects to play with 
  immediately without the need to jump through hoops, making it appealing. 
  The software implementation is executed well.}
\reply{Thank you.}

\pointRaised{Comment 2}{There are, however, a few inconsistencies in the implementation and some 
  issues with specific sections in the paper. In the following both issues 
  will be addressed sequentially by their occurrence in the paper.}
\reply{TBD}

\pointRaised{Comment 3}{p.4 illustrates the use of messages. The class implements list-like 
  access via \texttt{[[} and \$, but strangely \texttt{names()} return NULL and \texttt{length() }
  doesn't correspond to the number of fields leading to startling results like
the following:}

\begin{verbatim}
 > p
[1] "message of type 'tutorial.Person' with 2 fields set"
 > length(p)
[1] 2
 > p[[3]]
[1] ""
\end{verbatim}
\reply{TBD}

\pointRaised{Comment 3 cont.}{The inconsistencies get even more bizarre with descriptors (p.9):}

\begin{verbatim}
 > tutorial.Person$email
[1] "descriptor for field 'email' of type 'tutorial.Person' "
 > tutorial.Person[["email"]]
Error in tutorial.Person[["email"]] : this S4 class is not subsettable
 > names(tutorial.Person)
NULL
 > length(tutorial.Person)
[1] 1
\end{verbatim}
\reply{TBD}

\pointRaised{Comment 3 cont.}{It appears that there is no way to find out the fields of a descriptor 
  directly (although the low-level object methods seem to be exposed as 
  \texttt{\$field\_count()} and \texttt{\$fields()} - but that seems extremely cumbersome). 
  Again, implementing names() and subsetting may help here.}
\reply{TBD}

\pointRaised{Comment 4}{Another inconsistency concerns the \texttt{as.list()} method which by design 
  coerces objects to lists (see \texttt{?as.list}), but the implementation for 
  EnumDescriptor breaks that contract and returns a vector instead:}

\begin{verbatim}
 > is.list(as.list(tutorial.Person$PhoneType))
[1] FALSE
 > str(as.list(tutorial.Person$PhoneType))
  Named int [1:3] 0 1 2
  - attr(*, "names")= chr [1:3] "MOBILE" "HOME" "WORK"
\end{verbatim}

\pointRaised{Comment 4 cont}{As with the other interfaces, names() returns NULL so it is again quite
  difficult to perform even simple operations such as finding out the 
  values. It may be natural use some of the standard methods like names(), 
  levels() or similar. As with the previous cases, the lack of [[ support
  makes it impossible to map named enum values to codes and vice-versa.}
\reply{TBD}

\pointRaised{Comment 5}{In general, the package would benefit from one pass of checks to assess
  the consistency of the API. Since the authors intend direct interaction
  with the objects via basic standard R methods, the classes should behave 
  consistently.}
\reply{TBD}

\pointRaised{Comment 6}{Finally, most classes implement coercion to characters, which is not 
  mentioned and is not quite intuitive for some objects. For example, one
  may think that as.character() on a file descriptor returns let's say the 
  filename, but we get:}

\begin{verbatim}
 > cat(as.character(tutorial.Person$fileDescriptor()))
syntax = "proto2";

package tutorial;

option java_package = "com.example.tutorial";
option java_outer_classname = "AddressBookProtos";
[...]
\end{verbatim}
\reply{TBD}

\pointRaised{Comment 7}{It is not necessary clear what java\_package has to do with a file 
  descriptor in R. Depending on the intention here, it may be useful to 
  explain this feature.
}
\reply{TBD}

\subsubsection*{Other comments:}

\pointRaised{Comment 8}{p.17: "does not support ... function, language or environment. Such 
  objects have no native equivalent type in Protocol Buffers, and have 
  little meaning outside the context or R"
  That is certainly false. Native mirror of environments are hash tables - 
  a very useful type indeed. Language objects are just lists, so there is
  no reason to not include them - they can be useful to store expressions
  that may not be necessary specific to R. Further on p. 18 your run into
  the same problem that could be fixed so easily.}
\reply{TBD}

\pointRaised{Comment 9}{The examples in sections 7 and 8 are somewhat weak. It does not seem 
  clear why one would wish to unleash the power of PB just to transfer 
  breaks and counts for plotting - even a simple ASCII file would do that
  just fine. The main point in the example is presumably that there are 
  code generation methods for Hadoop based on PB IDL such that Hadoop can
  be made aware of the data types, thus making a histogram a proper record 
  that won't be split, can be combined etc. -- yet that is not mentioned 
  nor a way presented how that can be leveraged in practice. The Python 
  example code simply uses a static example with constants to simulate the 
  output of a reducer so it doesn't illustrate the point - the reader is 
  left confused why something as trivial would require PB while a savvy 
  reader is not able to replicate the illustrated process. Possibly 
  explaining the benefits and providing more details on how one would 
  write such a job would make it much more relevant.}
\reply{TBD}

\pointRaised{Comment 10}{Section 8 is not very well motivated. It is much easier to use other 
  formats for HTTP exchange - JSON is probably the most popular, but even
  CSV works in simple settings. PB is a much less common standard. The 
  main advantage of PB is the performance over the alternatives, but HTTP
  services are not necessarily known for their high-throughput so why one
  would sacrifice interoperability by using PB (they are still more hassle 
  and require special installations)? It would be useful if the reason 
  could be made explicit here or a better example chosen.}
\reply{TBD}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2014-12-15 papers/jss/response-to-reviewers.tex

\subsubsection*{More big picture, less details}

\pointRaised{Comment 2}{Overall, I think the paper provides too much detail on
  relatively unimportant topics and not enough on the reasoning behind
  important design decisions. I think you could comfortably reduce the paper
  by 5-10 pages, referring the interested reader to the documentation for
  more detail.}
\reply{The paper was rewritten throughout and is now much tighter at just 23 pages.}

\pointRaised{Comment 3}{I'd recommend shrinking section 3 to ~2 pages, and removing the
  subheadings. This section should quickly orient the reader to the
  RProtobuf API so they understand the big picture before learning more
  details in the subsequent sections. I'd recommend picking one OO style
  and sticking to it in this section - two is confusing.}
\reply{We followed this recommendation and reduced section 3 to about 2 1/2 pages.}

\pointRaised{Comment 3}{Section 4 dives into the details without giving a good overview and
  motivation. Why use S4 and not RC? How are the objects made mutable?
  Why do you provide both generic function and message passing OO
  styles? What does \$ do in this context? What the heck is a
  pseudo-method? Spend more time on those big issues rather than
  describing each class in detail. Reduce class descriptions to a
  bulleted list giving a high-level overview, then encourage the reader
  to refer to the documentation for further details. Similarly, Tables
  3-5 belong in the documentation, not in a vignette/paper.}
\reply{Done. RProtoBuf was designed and implemented before RC were
  available, and this is noted in a footnote now.  Explanation of how
  they are made mutable haas been added.  Better explanation of the
  two styles and '\$' as been added, while no longer using the
  confusing term
  'pseudo-method' anywhere.  Moved Tables 3-5 into the documentation
  and out of the paper, as suggested.}

\pointRaised{Comment 4}{Section 7 is weak. I think the important message is that RProtobuf is
  being used in practice at large scale for for large data, and is
  useful for communicating between R and Python. How can you make that
  message stronger while avoiding (for the purposes of this paper) the
  relatively unimportant details of the map-reduce setup?}
\reply{Done.  Rewritten with more motivation taking into account this feedback.}

@@

\subsubsection*{R to/from Protobuf translation}

\pointRaised{Comment 5}{The discussion of R to/from Protobuf could be improved. Table 9 would be
  much simpler if instead of Message, you provided a "vectorised"
  Messages class (this would also make the interface more consistent and
  hence the package easier to use).}
\reply{This is an area for future work and is a space explored in
  another package called Motobuf by other authors.}

\pointRaised{Comment 6}{Along these lines, I think it would make sense to combine sections 5
  and 6 and discuss translation challenges in both direction
  simultaneously. At the minimum, add the equivalent for Table 9 that
  shows how important R classes are converted to their protobuf
  equivalents.}

\reply{We have updated these sections to make it clearer that the main
  distinction is between schema-based datastructures (section 5) and
  schema-less use where a catch-all .proto is used (section 6).
  Neither section is meant to focus on only a single direction of the
  conversion, but how conversion works when you have a schema or not.
  How important R classes are converted to their protobuf equivalents
  isn't super useful as a C++, Java, or Python program is unlikely to
  want to read in an R data.frame exactly as it is defined.  Much more
  likely is an application-specific message format is defined between the
  two services, such as the HistogramTools example in the next section.
  Much more detail has been added to an interesting part of section 6 --
  which datasets exactly are better served with RProtoBuf than
  base::serialize and why?}

@@

\pointRaised{Comment 7}{You should discuss how missing values are handled for strings and
  integers, and why enums are not equivalent to factors. I think you
  could make explicit how coercion of factors, dates, times and matrices
  occurs, and the implications of this on sharing data structures
  between programming languages. For example, how do you share date/time
  data between R and python using RProtoBuf?}
\reply{TBD}

\pointRaised{Comment 8}{Table 10 is dying to be a plot, and a natural companion would be to
  show how long it takes to serialise data frames using both RProtoBuf
  and R's native serialisation. Is there a performance penalty to using
  protobufs?}
\reply{Table 10 has been replaced with a plot, the outliers are
  labeled, and the text now includes some interesting explanation
  about the outliers.  Page 4 explains that the R implementation of
  protocol buffers uses reflection to make operations slower but makes
  it more convenient for interactive data analysis.  None of the
  built-in datasets are large enough for performance to really come up
  as an issue, and for any serialization method examples could be
  found that significantly favor one over another, so we don't think
  there will be benefit to adding anything here.
}

@@

\subsubsection*{RObjectTables magic}

\pointRaised{Comment 9}{The use of RObjectTables magic makes me uneasy. It doesn't seem like a
  good fit for an infrastructure package and it's not clear what
  advantages it has over explicitly loading a protobuf definition into
  an object.}
\reply{More information about the advantages and disadvantages of this
  approach have been added.}

\pointRaised{Comment 10}{Using global state makes understanding code much harder. In Table 1,
  it's not obvious where \texttt{tutorial.Person} comes from. Is it loaded by
  default by RProtobuf? This need some explanation. In Section 7, what
  does \texttt{readProtoFiles()} do? Why does \texttt{RProtobuf} need to be attached
  as well as \texttt{HistogramTools}? This needs more explanation, and a
  comment on the implications of this approach on CRAN packages and
  namespaces.}
\reply{We followed this recommendation and added explanation for how
\texttt{tutorial.Person} is loaded, specifically : \emph{A small number of message types are imported when the
package is first loaded, including the tutorial.Person type we saw in
the last section.}  We removed the superfluous attach of \texttt{RProtoBuf}.}

\pointRaised{Comment 11}{
  I'd prefer you eliminate this magic from the magic, but failing that,
  you need a good explanation of why.}
\reply{We've added more explanation about this.}

@@

\subsubsection*{Code comments}

\pointRaised{Comment 12}{Using \texttt{file.create()} to determine the absolute path seems like a bad idea.}
\reply{We followed this recommendation and removed two instances of
  \texttt{file.create()} for this purpose with calls to
  \texttt{normalizePath} with \texttt{mustWork=FALSE}.}

### 2015-04-13 papers/jss/article.Rnw
today's changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@940 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\begin{itemize}
\item The functional dispatch mechanism of the the form
  \verb|method(object, arguments)| (common to \proglang{R}), and
\item The message passing object-oriented notation of the form
  \verb|object$method(arguments)|.
\end{itemize}

@@

<<>>=
f <- tutorial.Person$fileDescriptor()
f
f$Person
@

@@

\begin{table}[h]
\centering
\begin{small}
\begin{tabular}{lp{5cm}p{5cm}}
\toprule
Field type & \proglang{R} type (non repeated) & \proglang{R} type (repeated) \\
\cmidrule(r){2-3}
double	& \code{double} vector & \code{double} vector \\
float	& \code{double} vector & \code{double} vector \\[3mm]
uint32	  & \code{double} vector & \code{double} vector \\
fixed32	  & \code{double} vector & \code{double} vector \\[3mm]
int32	  & \code{integer} vector & \code{integer} vector \\
sint32	  & \code{integer} vector & \code{integer} vector \\
sfixed32  & \code{integer} vector & \code{integer} vector \\[3mm]
int64	  & \code{integer} or \code{character}
vector    & \code{integer} or \code{character} vector \\
uint64	  & \code{integer} or \code{character} vector & \code{integer} or \code{character} vector \\
sint64	  & \code{integer} or \code{character} vector & \code{integer} or \code{character} vector \\
fixed64	  & \code{integer} or \code{character} vector & \code{integer} or \code{character} vector \\
sfixed64  & \code{integer} or \code{character} vector & \code{integer} or \code{character} vector \\[3mm]
bool	& \code{logical} vector & \code{logical} vector \\[3mm]
string	& \code{character} vector & \code{character} vector \\
bytes	& \code{character} vector & \code{character} vector \\[3mm]
enum & \code{integer} vector & \code{integer} vector \\[3mm]
message & \code{S4} object of class \code{Message} & \code{list} of \code{S4} objects of class \code{Message} \\
\bottomrule
\end{tabular}
\end{small}
\caption{\label{table-get-types}Correspondence between field type and
  \proglang{R} type retrieved by the extractors. Note that \proglang{R} lacks native
  64-bit integers, so the \code{RProtoBuf.int64AsString} option is
  available to return large integers as characters to avoid losing
  precision.  This option is described in Section~\ref{sec:int64}.}
\end{table}

@@

% latex table generated in R 3.0.2 by xtable 1.7-0 package
% Wed Nov 26 15:31:30 2014
%\begin{table}[ht]
%\begin{center}
\begin{tabular}{rlrrrrr}
  \toprule
  Data Set & object.size & \multicolumn{2}{c}{\proglang{R} Serialization} &
  \multicolumn{2}{c}{RProtoBuf Serialization} \\
  & & default & gzipped & default & gzipped \\
  \cmidrule(r){2-6}
 crimtab & 7,936 & 4,641 (41.5\%) & 713 (91.0\%) & 1,655 (79.2\%) & 576 (92.7\%)\\
 airquality & 5,496 & 4,551 (17.2\%) & 1,241 (77.4\%) & 2,874 (47.7\%) & 1,294 (76.5\%)\\
 faithful & 5,136 & 4,543 (11.5\%) & 1,339 (73.9\%) & 4,936 (3.9\%) & 1,776 (65.5\%)\\
   \hline
 All & 605,256 & 461,667 (24\%) & 138,937 (77\%) & 435,360 (28\%) & 142,134 (77\%)\\
\hline
\end{tabular}
\end{center}
\caption{(Top) Relative space savings of Protocol Buffers and native \proglang{R} serialization over the raw object sizes of each of the \Sexpr{n} data sets in the \pkg{datasets} package. Points to the left of the dashed $y=x$ line represent datasets that are more efficiently encoded with Protocol Buffers. (Bottom) Absolute space savings of three outlier datasets and the aggregate performance of all datasets.}
\label{fig:compression}
\end{figure}

>>>>

\begin{itemize}
\item The functional dispatch mechanism of the the form
  \verb|method(object, arguments)| (common to \proglang{R}).
\item The message passing object-oriented notation of the form
  \verb|object$method(arguments)|.
\end{itemize}

@@

% < < > > =
% f <- tutorial.Person$fileDescriptor()
% f
% f$Person
% @

\begin{Schunk}
\begin{Sinput}
R> f <- tutorial.Person$fileDescriptor()
R> f
\end{Sinput}
\begin{Soutput}
file descriptor for package tutorial \
    (/usr/local/lib/R/site-library/RProtoBuf/proto/addressbook.proto)
\end{Soutput}
\begin{Sinput}
R> f$Person
\end{Sinput}
\begin{Soutput}
descriptor for type 'tutorial.Person' 
\end{Soutput}
\end{Schunk}

@@

\begin{table}[h]
\centering
\begin{small}
\begin{tabular}{lp{5cm}p{5.5cm}}
\toprule
Field type & \proglang{R} type (non repeated) & \proglang{R} type (repeated) \\
\cmidrule(r){2-3}
double	& \code{double} vector & \code{double} vector \\
float	& \code{double} vector & \code{double} vector \\[3mm]
uint32	  & \code{double} vector & \code{double} vector \\
fixed32	  & \code{double} vector & \code{double} vector \\[3mm]
int32	  & \code{integer} vector & \code{integer} vector \\
sint32	  & \code{integer} vector & \code{integer} vector \\
sfixed32  & \code{integer} vector & \code{integer} vector \\[3mm]
int64	  & \code{integer} or \code{character}
vector    & \code{integer} or \code{character} vector \\
uint64	  & \code{integer} or \code{character} vector & \code{integer} or \code{character} vector \\
sint64	  & \code{integer} or \code{character} vector & \code{integer} or \code{character} vector \\
fixed64	  & \code{integer} or \code{character} vector & \code{integer} or \code{character} vector \\
sfixed64  & \code{integer} or \code{character} vector & \code{integer} or \code{character} vector \\[3mm]
bool	& \code{logical} vector & \code{logical} vector \\[3mm]
string	& \code{character} vector & \code{character} vector \\
bytes	& \code{character} vector & \code{character} vector \\[3mm]
enum & \code{integer} vector & \code{integer} vector \\[3mm]
message & \code{S4} object of class \code{Message} & \code{list} of \code{S4} objects of class \code{Message} \\
\bottomrule
\end{tabular}
\end{small}
\caption{\label{table-get-types}Correspondence between field type and
  \proglang{R} type retrieved by the extractors. \proglang{R} lacks native
  64-bit integers, so the \code{RProtoBuf.int64AsString} option is
  available to return large integers as characters to avoid losing
  precision; see Section~\ref{sec:int64} below. 
  All but the \code{Message} type can be represented in vectors of one or
  more elements; for the latter a list is used.}
\end{table}

@@

% latex table generated in R 3.0.2 by xtable 1.7-0 package
% Wed Nov 26 15:31:30 2014
%\begin{table}[ht]
%\begin{center}
\begin{tabular}{rlrrrrr}
  \toprule
  Data set & object.size & \multicolumn{2}{c}{\proglang{R} serialization} &
  \multicolumn{2}{c}{RProtoBuf serialization} \\
  & & default & gzipped & default & gzipped \\
  \cmidrule(r){2-6}
 crimtab & 7,936 & 4,641 (41.5\%) & 713 (91.0\%) & 1,655 (79.2\%) & 576 (92.7\%)\\
 airquality & 5,496 & 4,551 (17.2\%) & 1,241 (77.4\%) & 2,874 (47.7\%) & 1,294 (76.5\%)\\
 faithful & 5,136 & 4,543 (11.5\%) & 1,339 (73.9\%) & 4,936 (3.9\%) & 1,776 (65.5\%)\\
   \hline
 All & 605,256 & 461,667 (24\%) & 138,937 (77\%) & 435,360 (28\%) & 142,134 (77\%)\\
\hline
\end{tabular}
\end{center}
\caption{(Top) Relative space savings of Protocol Buffers and native \proglang{R} serialization over the raw object sizes of each of the \Sexpr{n} data sets in the \pkg{datasets} package. Points to the left of the dashed $y=x$ line represent datasets that are more efficiently encoded with Protocol Buffers. (Bottom) Absolute space savings of three outlier datasets and the aggregate performance of all datasets.}
\label{fig:compression}
\end{figure}
